Thermodynamic Structure of the Solar Corona: Tomographic Reconstructions and MHD Modeling

### 2020-01-24 lloveras_2020.tex
modificaciones fede

\author[addressref={aff1,aff2},corref,email={albert@iafe.uba.ar}]{\inits{A.M.}\fnm{Alberto M.}~\lnm{V\'asquez}\orcid{0000-0003-3401-6409}}

\author[addressref={aff1,aff3},corref,email={federico@iafe.uba.ar}]{\inits{F.A.}\fnm{Federico A.}~\lnm{Nuevo}\orcid{0000-0003-2355-5853}}

\author[addressref={aff1,aff2},corref,email={cmaccormack@iafe.uba.ar}]{\inits{C.}\fnm{Cecilia}~\lnm{Mac Cormack}\orcid{0000-0003-1173-503X}}

\author[addressref={aff4},corref,email={nishthas@umich.edu}]{\inits{N.}\fnm{Nishtha}~\lnm{Sachdeva}\orcid{0000-0001-9114-6133}}

\author[addressref={aff4},corref,email={chipm@umich.edu}]{\inits{W.}\fnm{Ward}~\lnm{Manchester IV}\orcid{0000-0003-0472-9408}}

\author[addressref={aff4},corref,email={bartvand@umich.edu}]{\inits{B.}\fnm{Bartholomeus}~\lnm{Van der Holst}\orcid{0000-0001-5260-3944}}

\author[addressref={aff4},corref,email={rfrazin@umich.edu}]{\inits{R.A.}\fnm{Richard A.}~\lnm{Frazin}\orcid{0000-0002-0281-7677}}

@@

{To study the global coronal 3D structure, SRT proved to be a powerful tool. Based on solar rotation, allow instruments to compile a time series of images that cover the entire corona collecting data from all lines of sight.} In particular, differential emission measure tomography (DEMT), based on time-series of extreme ultraviolet (EUV) images, allows reconstruction of the 3D distribution of the differential emission measure (DEM) of the {quiet-Sun low corona at a global scale.} The technique was first developed by \citet{frazin_2009}, and first applied to the observational study of coronal structures by \citet{vasquez_2009}. Using solar rotation time series of EUV images taken in channels sensitive to different temperatures, the technique allows calculation of 3D maps of electron density and temperature in the heliocentric height range $1.0\,\mrsun$ to $1.2\,\mrsun$ as summarized in Section \ref{demt}. The reader may refer to the recent review by \citet{vasquez_2016} that summarizes the DEMT technique.

A combination of the DEMT and global magnetic models has provided useful insight into the 3D thermodynamical structure of the global quiet-Sun corona. In previous works, DEMT has been combined with a potential field source surface (PFSS) model  (\citet{huang_2012}, \citet{nuevo_2013}). Recently, \citet{lloveras_2017} combined DEMT with PFSS to study the thermodynamic of the last two solar minima along magnetic field lines and selecting different magnetic structures and \citet{maccormack_2017} developed a new DEMT product concerning the energetics of the inner solar corona throgh PFSS magnetic lines. In this article, for first time, we combined the DEMT reconstruction with the magnetic field of a MHD model, as explained in the next paragraph.

{The Alfv\'{e}n Wave Solar atmosphere Model (AWSoM) \diego{within the Space Weather Modeling Framework (SWMF)} is a three-dimensional (3D) physics-based, data-driven model extending from the upper chromosphere, into the solar corona upto 1 AU (and beyond) \citep{vander_2010, Van2014} and is constantly validated with observations as new improvements are implemented. DEMT results were used by \citet{jin_2012} and \citet{oran_2015} to validate AWSoM results finding an agreement within 50\% in density and electron temperature in the low corona. More recently, \citet{sachdeva_2019} compared the results of the latest version of AWSoM model with DEMT products in a globally fashion, obtaining an agreement within 20\%.
We used AWSoM results with two purposes. On the one hand to provide thermodynamic results to be contrasted with DEMT, and on the other hand to provide magnetohydrodynamic (MHD) coronal magnetic field model that is used to study thermodynamics along field lines.
}

@@

Section \ref{demt} and \ref{awsom} summarizes DEMT technique, and  AWSoM model respectively. Section \ref{trace} explains the method that is used to trace the thermodynamic results along the field lines of the magnetic model. Section \ref{energia} details the energy model and the loop integrated quantities. In Section \ref{demt_res} the quantitative comparative analysis of both rotations using DEMT is described and in Section \ref{awsom_res} AWSoM results are related to the one of DEMT. Finally in Section \ref{discu} summarizes and discusses the main conclusions of this analysis, and anticipates further planned work.

@@

{We carried out the 3D DEMT reconstruction of two target Carrington rotations (CR). We selected CR-2082, a deep minimum rotation between solar cycles (SC) 23 and 24, and CR-2208, a rotation during the early declining fase of SC 24. To study CR-2082 and CR-2208 data taken by the STEREO/EUVI-B and SDO/AIA instruments was used, respectively. The EUVI and AIA data were prepared using the latest processing tools and calibration corrections provided by their teams through the SolarSoft package. For this work, we introduced two improvements in the implementation of the DEMT technique, as described next.}

@@

In a first step, the time series of EUV images is used to solve a solar rotational tomography (SRT) problem, for each EUV band independently. As a result, the 3D distribution of the so called \emph{filter band emissivity} (FBE) is determined for each band separately. The FBE, {an emissivity-type quantity}, is defined as the wavelength integral of the coronal EUV spectral emissivity and the telescope's passband function of each EUV channel. Line-of-sight (LOS) integration of the FBE provides synthetic images that can be quantitatively compared to the real data in the time series. To find the FBE, the tomographic problem is posed a global optimization problem in which the quadratic norm of the difference between all {pairs of synthetic and real images is minimized.}

{In a second step, the FBE values obtained for all bands in each voxel {of the tomographic grid are used to constrain the determination of the LDEM which, as described in Section \ref{intro}, describes the temperature distribution of the plasma within the individual voxel.} Specifically, at each tomographic voxel $i$, the FBE of the band $k$ is related to the LDEM of the voxel according to}

@@

\noindent
{where $K$ is the number EUV bands, and $\TRF^{(k)}$ is the \emph{temperature response function} of the $j$-th detector. The TRFs are here computed based on the (known) channel’s passband times the coronal emissivity at that temperature (normalized by the squared electron density). The emissivity model used here is provided by the latest version of the CHIANTI atomic database and plasma emission model \citep{delzanna_2015,landi_2013}.}

Due to unresolved coronal dynamics, tomographic reconstructions exhibit negative values of the reconstructed FBE, or zero when the solution is constrained to positive values \citep{frazin_2000,frazin_2009}. These non-reconstructed voxels are indicated {in black color in the latitude-longitude (Carrington) maps of DEMT} reults in Section \ref{resu}.

The cadence of the data time-series is set to 6\,hr.

{In this work the cadence of the data time-series was set to 6\,hr and} data from three EUV bands were {used:} 171, 193 and 211 \AA \ in {the} case of AIA, and 171, 195 and 284 \AA \ bands in {the} case of EUVI. When using data from {three bands}, a Gaussian model for the LDEM is able to accurately predict the tomographic emissivities \citep{frazin_2009,nuevo_2015}. In each tomographic voxel the problem is then reduced to finding the values of the three free parameters of the Gaussian (centroid, standard deviation, and area) that allow to best reproduce the three tomographically reconstructed values of FBE in that voxel.

@@

We define next a measure of the accuracy of the LDEM model to predict the tomographic FBEs in each voxel, as

@@

Global maps of the observed photospheric magnetic field are essential for accurate modeling of the solar corona and are the primary input for AWSoM. The model use GONG synoptic map as input for CR-2082 (2009-04-05 - 2009-05-03) and ADAPT-GONG global magnetic field map for CR-2208 (2018-09-02 to 2018-09-29). ADAPT-GONG maps are unavailable for CR-2082. GONG provides synoptic full disk line-of-sight surface map of the radial magnetic field component. The unobserved polar fields in GONG are estimated by fitting a polynomial to neighboring observed fields. An improvement over these maps is provided by the ADAPT model \citep{Wor2000} which creates synchronic-synoptic maps that uses supergranulation, meridional circulation, flux-transport and differential rotation to evolve the magnetic field on the Sun for better estimates of the unobserved photospheric field \citep{Arg2010, Hen2012}. 
In AWSoM, the 2D radial photospheric field is converted to 3D magnetic field using the PFSS model which specifies the initial condition. We take the source surface to be at $2.5 \, \mrsun$ and the magnetic field is radial at this height. The radial component specifies the boundary condition.
%ADAPT-GONG maps have multiple realizations based on how the polar fields are filled. We use realization 5 for CR2208.
Based on \citet{Cha2011} the stochastic heating exponent is taken to be 0.21 and amplitude is 0.18 to account for energy partioning between electrons and protons. The Alfv\'{e}n wave energy boundary condition is given by the poynting flux (S$_{A}$) of the outgoing wave $(\frac{S_{A}}{B})_{\odot} = 1.0\times 10^{6}$ Wm$^{-2}$T$^{-1}$ and the correlation length of the Alfv\'{e}n waves is set to, $L_{\perp} \sqrt{B} = 1.5 \times 10^{5}$ m$\sqrt{T}$.

@@

\subsection{{Tracing Results Along Magnetic Fieldlines}}\label{trace}

@@

\noindent
where $h \equiv r - 1\,\mrsun$ is the coronal height measured from the photosphere. In the electron density fit, $\l\,[\mrsun]$ is the density scale height and $N_0\,[\cminvc]$ is the electron density at the footpoint ($h=0$) of the loop. In the electron temperature fit, $m\,[\MK/\mrsun]$ is the slope and $T_0\,[\MK]$ is the electron temperature at the footpoint of the loop. The slope $m$ is the radial gradient of the electron temperature along the loop, which we denote as $m = \dr\Tm$ hereafter, being $\dr\equiv\er\cdot\nabla$ the radial derivative operator, where $\er$ is the heliocentric radial unit vector.

@@

{The high temperature of the corona requires heating mechanisms to compensate for the the losses. While the vast majority of the existing literature on coronal heating focuses on ARs, some studies have been dedicated to the heating of quiet-Sun regions. In particular,  \citet{maccormack_2017} developed a new application of the DEMT technique to estimate the energy input flux required at the base of quiet-Sun coronal loops to mantain them stable. The technique is based on tracing the DEMT results along field lines of a global coronal magnetic model, just as described in the previous Section \ref{trace}.}

{Consider} a simple energy balance for each magnetic flux tube, {in which the dominating losses of} radiative power ($E_r$) and thermal conduction power ($E_c$) are compensated {by a} coronal heating power ($E_h$) \citep{aschwanden_2004}:

@@

\begin{equation}\label{Er}
E_r = \int\,\textrm{d}T\,\LDEM(T)\,\Lambda(T)
\end{equation}

@@

\begin{equation}\label{phi_r}
\phi_r = \left( \frac{B_0 \ B_L}{B_0 + B_L} \right) \int_{0}^{L} ds \ \frac{E_r(s)}{B(s)}
\end{equation}

\begin{equation}\label{phi_c}
\phi_c = \left( \frac{B_0 \ F_{c,L} - B_L \ F_{c,0}}{B_0 + B_L} \right) 
\end{equation}

@@

{Both target rotations are highly axisymmetric, \textit{i.e.} characterized by a high azymuthal symmetry. Rotation CR-2082} showed two {small} active regions (ARs), both near latitude $+30\mdeg$ and {around longitudes} $50\mdeg$ and $120\mdeg${, respectively}. {Rotation CR-2208} showed two {ARs, both near latitude} $+5\mdeg$ and {around longitudes $140\mdeg$ and $300\mdeg$, respectively.}

@@

>>>>

\author[addressref={aff1,aff2},email={albert@iafe.uba.ar}]{\inits{A.M.}\fnm{Alberto M.}~\lnm{V\'asquez}\orcid{0000-0003-3401-6409}}

\author[addressref={aff1,aff3},email={federico@iafe.uba.ar}]{\inits{F.A.}\fnm{Federico A.}~\lnm{Nuevo}\orcid{0000-0003-2355-5853}}

\author[addressref={aff1,aff2},email={cmaccormack@iafe.uba.ar}]{\inits{C.}\fnm{Cecilia}~\lnm{Mac Cormack}\orcid{0000-0003-1173-503X}}

\author[addressref={aff4},email={nishthas@umich.edu}]{\inits{N.}\fnm{Nishtha}~\lnm{Sachdeva}\orcid{0000-0001-9114-6133}}

\author[addressref={aff4},email={chipm@umich.edu}]{\inits{W.}\fnm{Ward}~\lnm{Manchester IV}\orcid{0000-0003-0472-9408}}

\author[addressref={aff4},email={bartvand@umich.edu}]{\inits{B.}\fnm{Bartholomeus}~\lnm{Van der Holst}\orcid{0000-0001-5260-3944}}

\author[addressref={aff4},email={rfrazin@umich.edu}]{\inits{R.A.}\fnm{Richard A.}~\lnm{Frazin}\orcid{0000-0002-0281-7677}}

@@

{To study the global coronal 3D structure, SRT proved to be a powerful tool. Based on solar rotation, allow instruments to compile a time series of images that cover the entire corona collecting data from all lines of sight.} In particular, differential emission measure tomography (DEMT), based on time-series of extreme ultraviolet (EUV) images, allows reconstruction of the 3D distribution of the differential emission measure (DEM) of the {quiet-Sun low corona at a global scale.} The technique was first developed by \citet{frazin_2009}, and first applied to the observational study of coronal structures by \citet{vasquez_2009}. Using solar rotation time series of EUV images taken in channels sensitive to different temperatures, the technique allows calculation of 3D maps of electron density and temperature in the heliocentric height range $1.0\,\mrsun$ to $1.2\,\mrsun$ as summarized in Section \ref{demt}. The reader may refer to the review by \citet{vasquez_2016} that describes the DEMT technique and summarizes the principal publications based on it.

A combination of the DEMT and global magnetic models has provided useful insight into the 3D thermodynamical structure of the global quiet-Sun corona. In previous works, DEMT has been combined with a potential field source surface (PFSS) model  (\citealt{huang_2012}, \citealt{nuevo_2013}). Recently, \citet{lloveras_2017} combined DEMT with PFSS to study the thermodynamic variables along different magnetic structure of the last two solar minima, and selecting different magnetic structures and \citet{maccormack_2017} developed a new DEMT product that estimates the energy input flux at the coronal base that is required to maintain stable the coronal loop structures represented by the PFSS field lines. In this article, for first time, we combined the DEMT reconstruction with the magnetic field of a magnetohydrodynamic (MHD) model, as explained in the next paragraph.

{The Alfv\'{e}n Wave Solar atmosphere Model (AWSoM) \diego{within the Space Weather Modeling Framework (SWMF)} is a three-dimensional (3D) physics-based, data-driven MHD model extending from the upper chromosphere, into the solar corona upto 1 AU (and beyond) \citep{vander_2010, Van2014} and is constantly validated with observations as new improvements are implemented. DEMT results were used by \citet{jin_2012} and \citet{oran_2015} to validate AWSoM results finding an agreement within 50\% in density and electron temperature in the low corona. More recently, \citet{sachdeva_2019} compared the results of the latest version of AWSoM model with DEMT products in a globally fashion, obtaining an agreement within 20\%.
We used AWSoM results with two purposes. On the one hand to provide thermodynamic results to be compared with those reconstructed by DEMT, and on the other hand to provide MHD coronal magnetic field model that is used to study thermodynamic variables along field lines.
}

@@

Section \ref{demt} and \ref{awsom} summarizes DEMT technique and AWSoM model respectively. Section \ref{trace} explains the method used to trace the thermodynamic results along the field lines of the magnetic model. Section \ref{energia} details the energy model and the loop integrated quantities. In Section \ref{demt_res} the quantitative comparative analysis of both rotations using DEMT is described and in Section \ref{awsom_res} AWSoM results are related to the one of DEMT. Finally in Section \ref{discu} summarizes and discusses the main conclusions of this analysis, and anticipates further planned work.

@@

{We carried out the 3D DEMT reconstruction of two target CR. We selected CR-2082, a deep minimum rotation between SCs 23 and 24, and CR-2208, a rotation during the early declining fase of SC 24. To study CR-2082 and CR-2208, as mentioned in Section \ref{intro}, data taken by the STEREO/EUVI-B and SDO/AIA instruments was used, respectively. The EUVI and AIA data were prepared using the latest processing tools and calibration corrections provided by their teams through the SolarSoft package. For this work, we introduced two improvements in the implementation of the DEMT technique, as described next.}

@@

In a first step, the time series of EUV images is used to solve a solar SRT problem, for each EUV band independently. As a result, the 3D distribution of the so called \emph{filter band emissivity} (FBE) is determined for each band separately. The FBE, {an emissivity-type quantity}, is defined as the wavelength integral of the coronal EUV spectral emissivity and the telescope's passband function of each EUV channel. Line-of-sight (LOS) integration of the FBE provides synthetic images that can be quantitatively compared to the real data in the time series. To find the FBE, the tomographic problem is posed a global optimization problem in which the quadratic norm of the difference between all {pairs of synthetic and real images is minimized.}

{In a second step, the FBE values obtained for all bands in each voxel {of the tomographic grid are used to constrain the determination of the local-DEM (LDEM) which, as described in Section \ref{intro}, describes the temperature distribution of the plasma within the individual voxel.} Specifically, at each tomographic voxel $i$, the FBE of the band $k$ is related to the LDEM of the voxel according to}

@@

\noindent
{where $K$ is the number EUV bands, and $\TRF^{(k)}$ is the \emph{temperature response function} of the $k$-th detector. In this work, the TRFs are computed based on the (known) channel’s passband times the coronal emissivity at that temperature (normalized by the squared electron density). The emissivity model used here is provided by the latest version of the CHIANTI atomic database and plasma emission model \citep{delzanna_2015,landi_2013}.}

Due to unresolved coronal dynamics, tomographic reconstructions exhibit negative values of the reconstructed FBE, or zero when the solution is constrained to positive values \citep{frazin_2000,frazin_2009}. These non-reconstructed voxels are indicated {in black color in the latitude-longitude (Carrington) maps of DEMT} results in Section \ref{resu}.

{In this work the cadence of the data time-series was set to 6\,hr and} data from three EUV bands were {used:} 171, 193 and 211 \AA \ in {the} case of AIA, and 171, 195 and 284 \AA \ bands in {the} case of EUVI. When using data from {three bands}, a Gaussian model for the LDEM is able to accurately predict the FBEs \citep{frazin_2009,nuevo_2015}. In each tomographic voxel the problem is then reduced to finding the values of the three free parameters of the Gaussian (centroid, standard deviation, and area) that allow to best reproduce the three tomographically reconstructed values of FBE in that voxel.

@@

Next, we define a measure of the accuracy of the LDEM model to predict the tomographic FBEs in each voxel, as

@@

Global maps of the observed photospheric magnetic field are essential for accurate modeling of the solar corona and are the primary input for AWSoM. The model use GONG synoptic map as input for CR-2082 and ADAPT-GONG global magnetic field map for CR-2208. ADAPT-GONG maps are unavailable for CR-2082. GONG provides synoptic full disk line-of-sight surface map of the radial magnetic field component. The unobserved polar fields in GONG are estimated by fitting a polynomial to neighboring observed fields. An improvement over these maps is provided by the ADAPT model \citep{Wor2000} which creates synchronic-synoptic maps that uses supergranulation, meridional circulation, flux-transport and differential rotation to evolve the magnetic field on the Sun for better estimates of the unobserved photospheric field \citep{Arg2010, Hen2012}. 
In AWSoM, the 2D radial photospheric field is converted to 3D magnetic field using the PFSS model which specifies the initial condition. We take the source surface to be at $2.5 \, \mrsun$ and the magnetic field is radial at this height. The radial component specifies the boundary condition.
%ADAPT-GONG maps have multiple realizations based on how the polar fields are filled. We use realization 5 for CR2208.
Based on \citet{Cha2011} the stochastic heating exponent is taken to be 0.21 and amplitude is 0.18 to account for energy partioning between electrons and protons. The Alfv\'{e}n wave energy boundary condition is given by the poynting flux (S$_{A}$) of the outgoing wave $(\frac{S_{A}}{B})_{\odot} = 1.0\times 10^{6}$ Wm$^{-2}$T$^{-1}$ and the correlation length of the Alfv\'{e}n waves is set to, $L_{\perp} \sqrt{B} = 1.5 \times 10^{5}$ m$\sqrt{T}$.

@@

\subsection{{Tracing Results Along Magnetic Field lines}}\label{trace}

@@

\noindent
where $h \equiv r - 1\,\mrsun$ is the coronal height measured from the photosphere. In the electron density fit, $\l\,[\mrsun]$ is the density scale height and $N_0\,[\cminvc]$ is the electron density at the footpoint ($h=0$) of the loop. In the electron temperature fit, $m\,[\MK/\mrsun]$ is the slope and $T_0\,[\MK]$ is the electron temperature at the footpoint of the loop. The slope $m$ estimates the radial gradient of the electron temperature along the loop, which we denote as $m = \dr\Tm$ hereafter, being $\dr\equiv\er\cdot\nabla$ the radial derivative operator, where $\er$ is the heliocentric radial unit vector.

@@

{The high temperature of the corona requires heating mechanisms to compensate for the energy losses. While the vast majority of the existing literature on coronal heating focuses on active regions (ARs), some studies have been dedicated to the heating of quiet-Sun regions. In particular,  \citet{maccormack_2017} developed a new application of the DEMT technique to estimate the energy input flux required at the base of quiet-Sun coronal loops to mantain them stable. The technique is based on tracing the DEMT results along field lines of a global coronal magnetic model, just as described in Section \ref{trace}.}

{Consider} a static energy balance for each magnetic flux tube, {in which the dominating losses of} radiative power ($E_r$) and thermal conduction power ($E_c$) are compensated {by a} coronal heating power ($E_h$) \citep{aschwanden_2004}:

@@

\begin{equation}\label{Er}
E_r = \int\,\textrm{d}T\,\LDEM(T)\,\Lambda(T),
\end{equation}

@@

\begin{equation}\label{phi_r}
\phi_r = \left( \frac{B_0 \ B_L}{B_0 + B_L} \right) \int_{0}^{L} ds \ \frac{E_r(s)}{B(s)},
\end{equation}

\begin{equation}\label{phi_c}
\phi_c = \left( \frac{B_0 \ F_{c,L} - B_L \ F_{c,0}}{B_0 + B_L} \right) ,
\end{equation}

@@

{Both target rotations are highly axisymmetric, \textit{i.e.} characterized by a high azymuthal symmetry. Rotation CR-2082} showed two {small} ARs, both near latitude $+30\mdeg$ and {around longitudes} $50\mdeg$ and $120\mdeg${, respectively}. {Rotation CR-2208} showed two {ARs, both near latitude} $+5\mdeg$ and {around longitudes $140\mdeg$ and $300\mdeg$, respectively.}

@@

To develop a validation of the AWSoM model described in Section \ref{awsom} with the DEMT technique explained in Section \ref{demt}, we perform the AWSoM model to the same two target rotations shown in the previous section.
